---
title: "Spatial upscaling Exercise - Final Report"
author: "Manuel Kunz"
date: "2024-01-02"
bibliography: "references.bib"
output: html_document
codefolding: hide
---

## 1 Literature

@Ludwig2023

## 2 Random cross-validation
### Used model

### Results 
Root mean square error (RSME) and $R^2$ across cross-validation folds
```{r, echo= FALSE, results = 'asis', message=FALSE, warning=FALSE}
library(knitr)
results_foldwise <- readRDS(paste0(here::here(),"./data/results_foldwise.rds"))

kable(results_foldwise, caption = "Table 1: RSME and $R^2$ across cross-validation folds")
```
Model evaluation
```{r, model evaluation, message=FALSE, warning=FALSE, echo=FALSE}
plot_random_cv <- readRDS(paste0(here::here(),"./data/plots_cv_results.rds"))
plot_random_cv
```


## 3 Spatial cross-validation

### 3.1 Observations regarding the locations of the data points


```{r, map data points, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 2: Map showing the spatial distribution of the original data points (red)"}
map_datapoints <- readRDS(paste0(here::here(),"./data/map_datapoints.rds"))
map_datapoints
```





### 3.2 Performing a spatial cross-validation
To perform a spatial cross-validation, I divided the data points used in Chapter 3.1 into 5 groups based on their location (lat and lon coordinates). This was done using the kmeans function from the R {stats} package.

```{r, code kmeans, message=FALSE, warning=FALSE}
# load prepared data
dfs <- readRDS(here::here("./data/dfs.rds"))

# cluster the data
clusters <- kmeans(
  dfs[,2:3],    # column 2 contains lon and column contains lat data
  centers = 5   # 5 clusters should be created
)

# add the create clusters to the original data frame
dfs$cluster <- clusters$cluster
```

```{r, map data points clustered, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 3: Map showing the original data points (red) and their classification to the 5 created clusters (see legend)"}
     
map_datapoints_clustered <- readRDS(paste0(here::here(),"./data/map_datapoints_clustered.rds"))
map_datapoints_clustered
      
```

### 3.3 Distribution of leaf N by cluster
Figure 4 shows the distribution of leaf N by cluster using density plots. The x and y axes are fixed to the same ranges, so that the five plots can be compared.

```{r, density plots, echo = FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 4: Leaf N density plots for every cluster"}
density_plots <- readRDS(paste0(here::here(),"./data/density_plots.rds"))
density_plots
```


### 3.4 RSME and $R^2$ results



### 3.5 Comparison random and spatial cross validation

## 4 Environmental cross-validation
```{r, code kmeans env cv, message=FALSE, warning=FALSE}
# load prepared data
dfs <- readRDS(here::here("./data/dfs.rds"))

# cluster the data
clusters <- kmeans(
  dfs[,5:6],    # column 2 contains lon and column contains lat data
  centers = 5   # 5 clusters should be created
)

# add the create clusters to the original data frame
dfs$cluster <- clusters$cluster
```

```{r, map data points clustered env cv, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 3: Map showing the original data points (red) and their classification to the 5 created clusters (see legend)"}
     
map_datapoints_clustered_env <- readRDS(paste0(here::here(),"./data/map_datapoints_clustered_env.rds"))
map_datapoints_clustered_env
      
```

### 3.3 Distribution of leaf N by cluster
Figure 4 shows the distribution of leaf N by cluster using density plots. The x and y axes are fixed to the same ranges, so that the five plots can be compared.

```{r, density plots env cv, echo = FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 4: Leaf N density plots for every cluster"}
density_plots_env <- readRDS(paste0(here::here(),"./data/density_plots_env.rds"))
density_plots_env
```





## 5 References
