---
title: "Spatial upscaling Exercise - Final Report"
author: "Manuel Kunz"
date: "2024-01-02"
bibliography: "references.bib"
output: html_document
---

## 1 Literature

@Ludwig2023

## 2 Random cross-validation
### Used model

### Results 
Root mean square error (RSME) and $R^2$ across cross-validation folds
```{r, echo= FALSE, results = 'asis', message=FALSE, warning=FALSE}
library(knitr)
results_foldwise <- readRDS(paste0(here::here(),"./data/results_foldwise.rds"))

kable(results_foldwise, caption = "Table 1: RSME and $R^2$ across cross-validation folds")
```
Model evaluation
```{r, model evaluation, message=FALSE, warning=FALSE, echo=FALSE}
plot_random_cv <- readRDS(paste0(here::here(),"./data/plots_cv_results.rds"))
plot_random_cv
```


## 3 Spatial cross-validation

### 3.1 Observations regarding the locations of the data points


```{r, map data points, echo=FALSE, message=FALSE, warning=FALSE}
map_datapoints <- readRDS(paste0(here::here(),"./data/map_datapoints.rds"))
map_datapoints
```

```{r, map data points clustered, echo=FALSE, message=FALSE, warning=FALSE}
map_datapoints_clustered <- readRDS(paste0(here::here(),"./data/map_datapoints_clustered.rds"))
map_datapoints_clustered
```



### 3.2 Performing a spatial cross-validation


### 3.3 Distribution of leaf N by cluster



### 3.4 RSME and $R^2$ results



### 3.5 Comparison random and spatial cross validation

## 4 Environmental cross-validation






## 5 References
